Should I Airbnb My Property?

### 2017-03-09 abb_Writeup.Rmd
Fixed maps and vizs

---
title: "Should I Airbnb My Property?"
author: "Andy Krause and Gideon Aschwanden"
date: "March 10, 2017"
output:
  pdf_document: default
  html_document: default
header-includes: \usepackage{setspace}\doublespacing
geometry: margin=3.3cm
bibliography: abb.bib
---

@@

To address these questions and to provide some guidance to potential and current investors and owners, we examine the short- and long-term residential rental market in Melbourne, Australia. Melbourne has a thriving short-term rental market with over 20,000 total properties listed on the Airbnb portal, the source of our short-term market data. We adopt a scenario-based approach that compares observed annual revenues for short-term rental (STR) properties with an estimate of their expected long-term rental revenues. Our results show that the feasibility of a greater return through a short term rental -- what we have termed ``Short Term Preference'' -- is dependent on property type, product, location and host policies.  In short, the answer to whether or not a owner or investor should opt for a short-term leasing strategy: it depends.

@@

The Other category only makes up a very small percentage of the properties in both the short and long term market; especially so in the long-term market (\autoref{typecomp}) .  Additionally, a large proportion of the 'Others' in the short-term data are Bed & Breakfast units, a use which doesn't fit within our research question. As a result, we remove the 'Others' from both the short- and the long-term rental data.

@@

@@

We begin by removing all properties that are missing bedroom or bathroom information or have a likely data error in these fields (more than 14 bedrooms or 10 bathrooms). For both short- and long-term rentals we find that most apartments have fewer than four bedrooms and houses fewer than five.  For both property types, the vast majority of properties have three or fewer bathrooms. We filter the data accordingly. Additionally, note that short term rentals consider studio apartments to be 0 bed units while the long term data considers these as 1 bedroom. We convert the 0 bed short term units to 1 bed units. A few short term properties indicate no baths, which is likely an error so we remove these observations from the data. Also, here we notice that the short-term data gives bathrooms in halves, while the long term data presents baths in whole numbers.  We round the baths in the short-term data up to whole integers.

@@

```{r plot_bb, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=4, fig.width=5.75, fig.cap='\\label{bbtype}Bed and Baths by Tenure'}

@@

After removing property level observations that are missing price data, we plot the distribution of the rates and rents.  In Figure 4 we see that most long term weekly rentals are greater than \$200 and less than \$1,000 per week while most median nightly rates are greater than \$50 and less than \$500 per night. We filter the property level data accordingly. We also filter the short term daily data to include only those observations with nightly rates between \$50 and \$500.

@@

```{r plot_suburbmap, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=4, fig.width=4, fig.cap='\\label{locmap}Melbourne Submarkets', fig.lp='submap'}

@@

It is likely that the short-term rental market follows a similar market hierarchy, however, we can imagine that a few different spatial features can influence the short-term market.  While there are many possible additional features, we consider two to be the most likely to influence rates or occupancy: 1) Proximity to beaches; and 2) Proximity to key tourist activities and events. Using the three broad submarkets from the long-term market (inner, middle and outer) we have created a five sub-market system as a starting point for spatially analyzing short and long term rentals in Melbourne:

@@

```{r table_corelist, results="asis", comment=FALSE, message=FALSE, echo=FALSE, fig.cap='\\label{coretable}List of Core Suburbs', fig.lp='coresuburbs'}

@@

Before assigning submarkets to the long term property we must remedy the fact that about 5% of the long term rental observations have missing latitude and longitude values.  Half of these do, however, have a latitude and longitude values for the centroid of the street that the property faces.  To retain as many observations as possible in these intial steps we apply the street centroid value to the observations with missing lat/long values.  For those without either set of lat/longs we remove them from the dataset at this point. After making this correction we then assign one of the five submarket designations based on the process described above.

```{r plot_locations, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=5.5, fig.height=9, fig.cap='\\label{pointmaps}Location Maps', fig.lp='locmaps'}

@@

@@

To better visualize the relative frequency of the submarkets in the two tenure types we have broken down the data by submarket by property type by tenure type. A number of interesting observations are illustrated here (\autoref{smcounts}). First, the vast majority of Airbnb units are apartments in city-core or city locations. As expected and as shown by the map, the long-term properties are more evenly spread across the metro region than the short-term properties.  Long-term houses are more peripheral than apartments, a result of urban development patterns and land economics. Beach properties are, relatively speaking, over-represented in the short term market, especially in apartments, while suburban and rural apartments, somewhat common in long-term markets, are nearly absent in the short-term market. Overall, the suburban and rural Airbnb markets are quite thin, later results focusing on these markets will need to be examined in light of this small sample size.

@@

For the long term data, over half of the observations fall outside of our considered time frame, reducing the sample size down to around 130,000 from 290,000.  Further consideration regarding property types, structural characteristics, location, etc. remove another 12,000 observations, leaving 120,092 in the final dataset.

```{r table_cleansummary, results="asis", comment=FALSE, message=FALSE, echo=FALSE, fig.cap='Data Cleaning Summary', fig.lp='cleansumm'}

@@

rownames(clean.table) <- NULL
clean.table$field <- c('Initial Count', 'Time', 'Property Type',
                       'Listing Type', 'Bedrooms', 'Baths', 'Bed/Bath Combo',
                       'Rates & Rents', 'Revenue', 'Location', 'Host Policy')
clean.table$strper <- 100*round(clean.table$strper, 2)
clean.table$ltrper <- 100*round(clean.table$ltrper, 2)
names(clean.table) <- c('Filter', 'Short Term', 'Cut', '% Cut',
                        'Long Term', 'Cut', '% Cut')
options(xtable.comment = FALSE)
print(xtable(clean.table, caption='Data Cleaning Summary\n',
       digits=0),
       include.rownames=FALSE,
       caption.placement='top')

@@

Second, many of the short term properties have a significant number of blocked days, days in which the property was not offered on Airbnb.  Here too, there is a problem with comparing observed revenues from a property with a 40\% block rate with a hypothetical year long revenue stream, especially if the property was leased out on other portal during those blocked days.  And, even if we knew that the blocked days were simply due to the host wanting to use the property themselves during that time, this deduction in revenue should not necessarily apply to the hypothetical investor in our examined scenario.  In short, in either case, the actual observed short-term revenue would be biased on the low end compared to what a revenue maximizing owner would be likely able to accomplish.

@@

As the observed calculation systematically under counts the revenue from properties that were not on the Airbnb portal for the entire year we disregard it going forward and consider the Actual revenue metric to be a better measure of actual market performance. We chart the distribution of the Actual ($REV_a$) and Potential ($REV_p$) short-term revenues against a distribution of the observed long term revenues for the long term properties. Revenues from the long-term rentals are far more consistent, with the vast majority totaling between \$12,000 to \$30,000 per year (\autoref{revdens}).  Actual short-term revenues range from neglible to highs of greater than \$80,000, with many of the properties showing very limited (less than \$10,000) revenue. The range of the distribution for the Potential short-term revenue is similar, however, the distribution itself is much flatter, with many more properties showing potential incomes in the \$20,000 to \$40,000 range.

@@

  stre <- str.data[,c('bedrooms', 'str.ext.revenue')]
  stre$tenure <- 'short-term actual   '
  names(stre)[2] <- 'revenue'
  strp <- str.data[,c('bedrooms', 'str.pot.revenue')]
  strp$tenure <- 'short-term potential   '
  names(strp)[2] <- 'revenue'
  ltrr <- ltr.data[,c('bedrooms', 'ltr.revenue')]
  ltrr$tenure <- 'long-term   '
  names(ltrr)[2] <- 'revenue'
  allr <- rbind(stre, strp, ltrr)

  ggplot(allr, aes(x=revenue, fill=tenure, color=tenure)) +
    geom_density(alpha=.5) +
    scale_fill_manual(values=c(ltr.col, abb.col[2], str.col)) +
    scale_color_manual(values=c(ltr.col, abb.col[2], str.col)) +
    ggtitle('Annual Revenues by Tenure Type') +
    xlab('Annual Revenue') +
    scale_x_continuous(breaks=c(seq(0,60000, by=20000)),
                       labels=c('$0', '$20,000', '$40,000', '$60,000'))+
    theme(legend.position='bottom',
          legend.title = element_blank(),
          plot.title = element_text(hjust = 0.5),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.title.y=element_blank()) +
    coord_cartesian(xlim=c(0, 70000))

@@

As the distribution of occupancy rates is not uniform the initial plot may be misleading as to the frequency of short-term rentals being more profitable than long term. In the right hand panel of \autoref{occpref} (right panel) we transform the x-axis to represent the percentile of the full distribution of occupancy rates. By doing so, we see that, in general, properties need to be in the upper 25th percentile of occupancy rates across the entire market in order to have a equal likelihood of being more profitable as a short-term rental than as a more traditional long term one.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=7, fig.height=3.5, fig.cap='\\label{occpref}Probability of Short Term Pref. vs Occ. Rate'}

@@

An SVM can be used as a classifier to delineate the boundaries between two (or more) classes of observations.  In this case, we delineate the boundary, \autoref{metrohm} (right panel) between situations where long-term rentals are more profitable (red) and where short term are (green).  In order to avoid an overly broad straight line delineation we use a polynomial fitting technique. Overall, from the SVM model we see that the short-term preference begins at about 325% occupancy rate, regardless of nightly rate.  As the occupancy rate increases, the required nighly rate decreases, down to near $100 once occupancies hit 65\% or greater.

@@

To gain a deeper understanding of the differences in short-term preference, we break the data down across a variety of characteristics and examine the preferability metrics in these submarkets. First, we bifurcate our data into houses and apartments.  Within the houses, 17.5\% of the properties are short-term preferable, while for the apartment that figure is nearly double at 30.5\%. Plotting the changes in short-term preferability against occupancy rates shows considerable similarity between houses and apartments (\autoref{smocc}).  The only measurable difference is in that for houses a property's occupancy rate needs to be 10 percentiles higher than apartments to show the same likelihood of short-term preference as an apartment. Or, in other words, a given house needs to outperform its peers by a greater extent than an apartment to achieve short term preference.

@@

```{r, fig.width=8, fig.height=4, echo=FALSE, fig.cap='\\label{smhm}Analysis of Percentile Rates: By Type'}

@@

 # Limit to the extrapolated pref type
  pref.table <- pref.table[pref.table$rev.type == 'Extrapolated', ]
  pref.table[,1] <- c('City-Core', 'City', 'Suburban', 'Rural', 'Beach')
  pref.table[,2] <- pref.table[,2]*100

  names(pref.table) <- c('Sub-Market', '% Short Term Preferable')
  options(xtable.comment = FALSE)
  print(xtable(pref.table[,1:2], caption='Geographic Submarket Preferability\n',
               digits=1),
               include.rownames=FALSE,
               caption.placement='top')

@@

Plotting out the nightly rate versus occupancy rate heat maps for all five geographic submarkets (\autoref{smhmps}) tells a similar story.  The 'City-Core' market offers the most potential for short term preference with suburban and city the least.  The rural market achieves very solid short-term preference for the most expensive homes, again likely those with special amenities or links to the wine region.  Beach properties do not perform as well as might be expected in the short term market, however, it should be kept in mind that the winter climate in Melbourne is not conducive to beach activity and therefore short-term beach properties may suffer during winter months.

@@

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='Logistic Model Results of Short Term Preference'}

@@

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='Potential Short Term Preference'}

@@

  spref.table <- spref.table[spref.table$rev.type == 'Potential', ]
  st.table <- rbind(tpref.table, spref.table)
  st.table[,2] <- st.table[,2]*100
  names(st.table) <- c('Sub-Market', '% Short Term Preferable')
  options(xtable.comment = FALSE)
  print(xtable(st.table[,1:2], 
               caption='Potential Revenue Preferability\n',
               digits=1),
               include.rownames=FALSE,
               caption.placement='top',
               hline.after=c(-1, 0, 2, 7, 7))

@@

  pqtl.hm.svm <- abbHeatMap(pot.data.ic,
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.pot.pref',
                   svm=T, 
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5]))
  pqtl.hm.svm <- pqtl.hm.svm + ylab('Nightly Rate (Qtl)')+
  annotate("text", x = 50, y = 50, label = "Potential", size=9)

@@

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.height=3.5, fig.cap='\\label{pothm}Heatmaps of Potential Short Term Preference'}

@@

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='Logistic Model Results of Potential Short Term Preference'}

@@

Overall, the choice of revenue type, Actual or Potential, does affect some of the conclusions to be derived from the results above.  As expected, under the Potential case there are more properties (50\% vs 27\%) that exhibit short-term preference.  Under the scenario, the possibilities of increasing revenue by moving from long to short term leasing appear much better.  Again, however, we have not attempted to measure or estimate how this hypothetical flood of properties onto the market (due to the elimation of blocked days) would impact overall market dynamics.  Suffice it to say that it would like drive down occupancies and maybe price, therefore leading to a decrease in short-term preference likelihood.

@@

host.type.bar <- ggplot(str.data, aes(host.type, fill=host.type)) +
  geom_bar() +
  scale_fill_manual(values=c(abb.col[2], abb.col[4], abb.col[7], abb.col[9]))+
  theme(legend.position='none') +
  ylab('Frequency\n') + 
  xlab('\nHost Type') +
  scale_x_discrete(labels=c('Profit','Sharer','Multi','Unk'))+
  #ggtitle('Short Term Host Types') + 
  theme(plot.title = element_text(hjust = 0.5))

@@

 # Limit to the extrapolated pref type
  hpref.table <- apref.table[apref.table$rev.type == 'Extrapolated', ]
  phpref.table <- apref.table[apref.table$rev.type == 'Potential', ]
  hpref.table <- hpref.table[,1:2]
  hpref.table[,3] <- phpref.table[,2]
  hpref.table[,2] <- hpref.table[,2]*100
  hpref.table[,3] <- hpref.table[,3]*100

  names(hpref.table) <- c('Host Type', '% STP - Actual', '% STP - Potential')
  options(xtable.comment = FALSE)
  print(xtable(hpref.table, 
               caption='Host Type Preferability\n',
               digits=1),
               include.rownames=FALSE,
               caption.placement='top')

@@

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='\\label{logpot}Logistic Model Results of Potential Short Term Preference'}

@@

In analyzing the short-term observations with estimated long-term rents we found that approximately one in four Airbnb properties brought in more revenue with its short-term leasing strategy than it would have with a long-term approach. As expected, this probability is not equal across space, property type or property size (bed and bath count).  Apartments are much more likely than houses to be more profitable as a short-term rental.  Core city properties are more likely than other properties to prefer a short-term strategy, as are smaller properties versus larger ones. For instance, 36\% of apartments in the city core generated more revenue on Airbnb than they likely would have on the long term market while only 8\% of suburban houses did. Looking at potential revenues -- through eliminating the impact of blocked days -- increases the overall likelihood of short-term preference considerably.

The nightly rates and total occupancies drive the profitability of short-term leasing.  As the nightly rate vs. occupancy rate heat maps and SVM plots (\autoref{metrohmp}, \ref{smhm}, \ref{smhmps}, \ref{pothm}) show, even with very high nightly rates, owners would need to achieve an occupancy rate of at least the 75th percentile in market (around 40\%) in order to achieve a better than 50\% probability of being more profitable in a short-term strategy than a long term one. A 40\% occupancy rate means bookings at least three days out of every week all year long. In Melbourne summers this may be an easy feat as tourist numbers swell, but doing so on a cold week in July likely proves more difficult. The vertical-ness of the divide between short- and long-term preference in the SVM plots suggests that occupancy rate is the most critical factor in reaching short-term preference in most markets.

The logistic regression models (Table 4, 6 and 8) than we estimate show that most of the finding from the univariate and bivariate analyses done earlier are validated .  Small apartments in the central city are the best bet for generating greater revenues through short term leasing.  We also test for the impact of host behavior on the probability of short term preference.  We find that more flexibility around the number of guests and the minimum stay increase the probability of short-term preference, while the reverse is true of cancellation policies (more strict equals higher short term preference probability).  As a host, it appears to be best to do what you can to attract bookings and then make it difficult for those bookings to be cancelled.

@@

There are a number of limitations to this research.  There are few shared property characterstics between the two datasets, which make cross-estimating rents and rates difficult, especially from the short-term market.  Part of the difficulty in estimating potential short-term rates for long-term properties is that the short term market appears to be heavily influenced by non-property specific factors such as the leasing policies (minimum stay, number of guests, etc.) themselves.  Future research should focus both on building better models to explain (and ultimately predict) short-term rates and occupancies as well as to understand the impact of policies such as strict cancellation or no pets on the short term market.  It is likely that this work is being done proprietarily by Airbnb.com itself and other complementary firms such as AirDNA.com, our data provider.

@@

<!-- ```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='Logistic Model Results of Short Term Preference', fig.lp='xxx'} -->

>>>>

---
title: "Should I Airbnb My Property?"
author: "Andy Krause and Gideon Aschwanden"
date: "March 10, 2017"
output:
  pdf_document: default
  html_document: default
header-includes: 
   - \usepackage{setspace}\doublespacing
   - \usepackage{caption}
   - \captionsetup[table]{skip=10pt}
geometry: margin=3.3cm
bibliography: abb.bib
---

@@

To address these questions and to provide some guidance to potential and current investors and owners, we examine the short- and long-term residential rental market in Melbourne, Australia. Melbourne has a thriving short-term rental market with over 20,000 total properties listed on the Airbnb portal, the source of our short-term market data. We adopt a scenario-based approach that compares observed annual revenues for short-term rental (STR) properties with an estimate of their expected long-term rental revenues. Our results show that the feasibility of a greater return through a short-term rental -- what we have termed ``Short-Term Preference'' -- is dependent on property type, product, location and host policies.  In short, the answer to whether or not a owner or investor should opt for a short-term leasing strategy: it depends.

@@

The Other category only makes up a very small percentage of the properties in both the short and long-term market; especially so in the long-term market (\autoref{typecomp}) .  Additionally, a large proportion of the 'Others' in the short-term data are Bed & Breakfast units, a use which doesn't fit within our research question. As a result, we remove the 'Others' from both the short- and the long-term rental data.

@@

  str.types <- str.types + xlab('')
  ltr.types <- ltr.types + xlab('')

@@

We begin by removing all properties that are missing bedroom or bathroom information or have a likely data error in these fields (more than 14 bedrooms or 10 bathrooms). For both short- and long-term rentals we find that most apartments have fewer than four bedrooms and houses fewer than five.  For both property types, the vast majority of properties have three or fewer bathrooms. We filter the data accordingly. Additionally, note that short-term rentals consider studio apartments to be 0 bed units while the long-term data considers these as 1 bedroom. We convert the 0 bed short-term units to 1 bed units. A few short-term properties indicate no baths, which is likely an error so we remove these observations from the data. Also, here we notice that the short-term data gives bathrooms in halves, while the long-term data presents baths in whole numbers.  We round the baths in the short-term data up to whole integers.

@@

```{r plot_bb, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=4, fig.width=5.75, fig.cap='\\label{bbtype}Bed and Baths by Tenure by Type'}

@@

After removing property level observations that are missing price data, we plot the distribution of the rates and rents.  In Figure 4 we see that most long-term weekly rentals are greater than \$200 and less than \$1,000 per week while most median nightly rates are greater than \$50 and less than \$500 per night. We filter the property level data accordingly. We also filter the short-term daily data to include only those observations with nightly rates between \$50 and \$500.

@@

```{r plot_suburbmap, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=3.5, fig.width=3.5, fig.cap='\\label{locmap}Melbourne Submarkets', fig.lp='submap'}

@@

It is likely that the short-term rental market follows a similar market hierarchy, however, we can imagine that a few different spatial features can influence the short-term market.  While there are many possible additional features, we consider two to be the most likely to influence rates or occupancy: 1) Proximity to beaches; and 2) Proximity to key tourist activities and events. Using the three broad submarkets from the long-term market (inner, middle and outer) we have created a five sub-market system as a starting point for spatially analyzing short and long-term rentals in Melbourne:

@@

```{r table_corelist, results="asis", comment=FALSE, message=FALSE, echo=FALSE, fig.cap='\\label{coretable}List of Core Suburbs'}

@@

Before assigning submarkets to the long-term property we must remedy the fact that about 5% of the long-term rental observations have missing latitude and longitude values.  Half of these do, however, have a latitude and longitude values for the centroid of the street that the property faces.  To retain as many observations as possible in these intial steps we apply the street centroid value to the observations with missing lat/long values.  For those without either set of lat/longs we remove them from the dataset at this point. After making this correction we then assign one of the five submarket designations based on the process described above.

```{r load_locations, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE}

@@

```{r plot_locations, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=5, fig.height=10, fig.cap='\\label{pointmaps}Location Maps', fig.lp='locmaps', fig.ext='png'}

@@

 To better visualize the relative frequency of the submarkets in the two tenure types we have broken down the data by submarket by property type by tenure type. A number of interesting observations are illustrated here (\autoref{smcounts}). First, the vast majority of Airbnb units are apartments in city-core or city locations. As expected and as shown by the map, the long-term properties are more evenly spread across the metro region than the short-term properties.  Long-term houses are more peripheral than apartments, a result of urban development patterns and land economics. Beach properties are, relatively speaking, over-represented in the short-term market, especially in apartments, while suburban and rural apartments, somewhat common in long-term markets, are nearly absent in the short-term market. Overall, the suburban and rural Airbnb markets are quite thin, later results focusing on these markets will need to be examined in light of this small sample size.

@@

For the long-term data, over half of the observations fall outside of our considered time frame, reducing the sample size down to around 130,000 from 290,000.  Further consideration regarding property types, structural characteristics, location, etc. remove another 12,000 observations, leaving 120,092 in the final dataset.

```{r table_cleansummary, results="asis", comment=FALSE, message=FALSE, echo=FALSE, fig.cap='Data Cleaning Summary'}

@@

rownames(clean.table) <- NULL
clean.table$field <- c('Initial Count', 'Time', 'Property Type',
                       'Listing Type', 'Bedrooms', 'Baths', 'Bed/Bath Combo',
                       'Rates & Rents', 'Revenue', 'Location', 'Host Policy')
clean.table$strper <- 100*round(clean.table$strper, 2)
clean.table$ltrper <- 100*round(clean.table$ltrper, 2)
names(clean.table) <- c('Filter', 'Short-term', 'Cut', '% Cut',
                        'Long-term', 'Cut', '% Cut')
options(xtable.comment = FALSE)
print(xtable(clean.table, caption='Data Cleaning Summary',
       digits=0),
       include.rownames=FALSE,
       caption.placement='top')

@@

Second, many of the short-term properties have a significant number of blocked days, days in which the property was not offered on Airbnb.  Here too, there is a problem with comparing observed revenues from a property with a 40\% block rate with a hypothetical year long revenue stream, especially if the property was leased out on other portal during those blocked days.  And, even if we knew that the blocked days were simply due to the host wanting to use the property themselves during that time, this deduction in revenue should not necessarily apply to the hypothetical investor in our examined scenario.  In short, in either case, the actual observed short-term revenue would be biased on the low end compared to what a revenue maximizing owner would be likely able to accomplish.

@@

As the observed calculation systematically under counts the revenue from properties that were not on the Airbnb portal for the entire year we disregard it going forward and consider the Actual revenue metric to be a better measure of actual market performance. We chart the distribution of the Actual ($REV_a$) and Potential ($REV_p$) short-term revenues against a distribution of the observed long-term revenues for the long-term properties. Revenues from the long-term rentals are far more consistent, with the vast majority totaling between \$12,000 to \$30,000 per year (\autoref{revdens}).  Actual short-term revenues range from neglible to highs of greater than \$80,000, with many of the properties showing very limited (less than \$10,000) revenue. The range of the distribution for the Potential short-term revenue is similar, however, the distribution itself is much flatter, with many more properties showing potential incomes in the \$20,000 to \$40,000 range.

@@

  stre <- str.data[,c('bedrooms', 'str.ext.revenue')]
  stre$tenure <- 'Short-Term (Actual)   '
  names(stre)[2] <- 'revenue'
  strp <- str.data[,c('bedrooms', 'str.pot.revenue')]
  strp$tenure <- 'Short-Term (potential)   '
  names(strp)[2] <- 'revenue'
  ltrr <- ltr.data[,c('bedrooms', 'ltr.revenue')]
  ltrr$tenure <- 'Long-Term   '
  names(ltrr)[2] <- 'revenue'
  allr <- rbind(stre, strp, ltrr)

  ggplot(allr, aes(x=revenue, fill=tenure, color=tenure)) +
    geom_density(alpha=.5) +
    scale_fill_manual(values=c(ltr.col, abb.col[2], str.col)) +
    scale_color_manual(values=c(ltr.col, abb.col[2], str.col)) +
    #ggtitle('Annual Revenues by Tenure Type') +
    xlab('\nAnnual Revenue') +
    scale_x_continuous(breaks=c(seq(0,60000, by=20000)),
                       labels=c('$0', '$20k', '$40k', '$60k'))+
    theme(legend.position='none',
          legend.title = element_blank(),
          plot.title = element_text(hjust = 0.5),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.title.y=element_blank()) +
    coord_cartesian(xlim=c(0, 70000)) +
    facet_wrap(~tenure)

@@

As the distribution of occupancy rates is not uniform the initial plot may be misleading as to the frequency of short-term rentals being more profitable than long-term. In the right hand panel of \autoref{occpref} (right panel) we transform the x-axis to represent the percentile of the full distribution of occupancy rates. By doing so, we see that, in general, properties need to be in the upper 25th percentile of occupancy rates across the entire market in order to have a equal likelihood of being more profitable as a short-term rental than as a more traditional long-term one.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=7, fig.height=3.5, fig.cap='\\label{occpref}Probability of Short-Term Preference vs Occupancy Rate'}

@@

An SVM can be used as a classifier to delineate the boundaries between two (or more) classes of observations.  In this case, we delineate the boundary, \autoref{metrohm} (right panel) between situations where long-term rentals are more profitable (red) and where short-term are (green).  In order to avoid an overly broad straight line delineation we use a polynomial fitting technique. Overall, from the SVM model we see that the short-term preference begins at about 325% occupancy rate, regardless of nightly rate.  As the occupancy rate increases, the required nighly rate decreases, down to near $100 once occupancies hit 65\% or greater.

@@

To gain a deeper understanding of the differences in short-term preference, we break the data down across a variety of characteristics and examine the preferability metrics in these submarkets. First, we bifurcate our data into houses and apartments.  Within the houses, 17.5\% of the properties are short-term preferable, while for the apartment that figure is nearly double at 30.5\%. Plotting the changes in short-term preferability against occupancy rates shows considerable similarity between houses and apartments (\autoref{smocc}).  The only measurable difference is in that for houses a property's occupancy rate needs to be 10 percentiles higher than apartments to show the same likelihood of short-term preference as an apartment. Or, in other words, a given house needs to outperform its peers by a greater extent than an apartment to achieve short-term preference.

@@

```{r, fig.width=8, fig.height=3.5, echo=FALSE, fig.cap='\\label{smhm}Analysis of Percentile Rates: By Type'}

@@

 # Limit to the extrapolated pref type
  pref.table <- pref.table[pref.table$rev.type == 'Extrapolated', ]
  pref.table[,1] <- c('City-Core', 'City', 'Suburban', 'Rural', 'Beach')
  pref.table[,2] <- paste0(sprintf('%.1f', pref.table[,2]*100), '%')

  names(pref.table) <- c('Sub-Market', 'Short-Term Preferable')
  options(xtable.comment = FALSE)
  print(xtable(pref.table[,1:2], caption='Geographic Submarket Preferability\n',
               digits=1,
               align='llr'),
               include.rownames=FALSE,
               caption.placement='top')

@@

Plotting out the nightly rate versus occupancy rate heat maps for all five geographic submarkets (\autoref{smhmps}) tells a similar story.  The 'City-Core' market offers the most potential for short-term preference with suburban and city the least.  The rural market achieves very solid short-term preference for the most expensive homes, again likely those with special amenities or links to the wine region.  Beach properties do not perform as well as might be expected in the short-term market, however, it should be kept in mind that the winter climate in Melbourne is not conducive to beach activity and therefore short-term beach properties may suffer during winter months.

@@

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='Logistic Model Results of Short-Term Preference'}

@@

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='Potential Short-Term Preference'}

@@

  spref.table <- spref.table[spref.table$rev.type == 'Potential', ]
  st.table <- rbind(tpref.table, spref.table)
  st.table[,2] <- paste0(sprintf('%.1f', st.table[,2]*100), '%') 
  names(st.table) <- c('Sub-Market', 'Short-Term Preferable')
  options(xtable.comment = FALSE)
  print(xtable(st.table[,1:2], 
               caption='Potential Revenue Preferability\n',
               digits=1, align='llr'),
               include.rownames=FALSE,
               caption.placement='top',
               hline.after=c(-1, 0, 2, 7, 7))

@@

  pqtl.hm.svm <- abbHeatMap(pot.data.ic,
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.pot.pref',
                   svm=T, 
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5]))
  pqtl.hm.svm <- pqtl.hm.svm + ylab('Nightly Rate (Pcntl)')+
    xlab('Occupancy Rate (Pcntl)') +
  annotate("text", x = 50, y = 50, label = "Potential", size=9)

@@

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.height=3.5, fig.cap='\\label{pothm}Heatmaps of Short-Term Preference'}

@@

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='Logistic Model Results of Potential Short-Term Preference'}

@@

Overall, the choice of revenue type, Actual or Potential, does affect some of the conclusions to be derived from the results above.  As expected, under the Potential case there are more properties (50\% vs 27\%) that exhibit short-term preference.  Under the scenario, the possibilities of increasing revenue by moving from long- to short-term leasing appear much better.  Again, however, we have not attempted to measure or estimate how this hypothetical flood of properties onto the market (due to the elimation of blocked days) would impact overall market dynamics.  Suffice it to say that it would like drive down occupancies and maybe price, therefore leading to a decrease in short-term preference likelihood.

@@

host.type.bar <- ggplot(str.data, aes(host.type, fill=host.type)) +
  geom_bar() +
  scale_fill_manual(values=c(abb.col[2], abb.col[4], abb.col[7], abb.col[9]))+
  theme(legend.position='none') +
  ylab('Frequency\n') + 
  xlab('\nHost Type') +
  scale_x_discrete(labels=c('Profit','Sharer','Multi','Unk'))+
  #ggtitle('short-term Host Types') + 
  theme(plot.title = element_text(hjust = 0.5))

@@

 # Limit to the extrapolated pref type
  hpref.table <- apref.table[apref.table$rev.type == 'Extrapolated', ]
  phpref.table <- apref.table[apref.table$rev.type == 'Potential', ]
  hpref.table <- hpref.table[,1:2]
  hpref.table[,3] <- phpref.table[,2]
  hpref.table[,2] <- paste0(sprintf('%.1f', hpref.table[,2]*100), '%')
  hpref.table[,3] <- paste0(sprintf('%.1f', hpref.table[,3]*100), '%')

  names(hpref.table) <- c('Host Type', 'STP (Actual)', 'STP (Potential)')
  options(xtable.comment = FALSE)
  print(xtable(hpref.table, 
               caption='Host Type Preferability\n',
               digits=1,
               align='llrr'),
               include.rownames=FALSE,
               caption.placement='top')

@@

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='\\label{logpot}Logistic Model Results of Potential Short-Term Preference'}

@@

In analyzing the short-term observations with estimated long-term rents we found that approximately one in four Airbnb properties brought in more revenue with its short-term leasing strategy than it would have with a long-term approach. As expected, this probability is not equal across space, property type or property size (bed and bath count).  Apartments are much more likely than houses to be more profitable as a short-term rental.  Core city properties are more likely than other properties to prefer a short-term strategy, as are smaller properties versus larger ones. For instance, 36\% of apartments in the city core generated more revenue on Airbnb than they likely would have on the long-term market while only 8\% of suburban houses did. Looking at potential revenues -- through eliminating the impact of blocked days -- increases the overall likelihood of short-term preference considerably.

The nightly rates and total occupancies drive the profitability of short-term leasing.  As the nightly rate vs. occupancy rate heat maps and SVM plots (\autoref{metrohmp}, \ref{smhm}, \ref{smhmps}, \ref{pothm}) show, even with very high nightly rates, owners would need to achieve an occupancy rate of at least the 75th percentile in market (around 40\%) in order to achieve a better than 50\% probability of being more profitable in a short-term strategy than a long-term one. A 40\% occupancy rate means bookings at least three days out of every week all year long. In Melbourne summers this may be an easy feat as tourist numbers swell, but doing so on a cold week in July likely proves more difficult. The vertical-ness of the divide between short- and long-term preference in the SVM plots suggests that occupancy rate is the most critical factor in reaching short-term preference in most markets.

The logistic regression models (Table 4, 6 and 8) than we estimate show that most of the finding from the univariate and bivariate analyses done earlier are validated .  Small apartments in the central city are the best bet for generating greater revenues through short-term leasing.  We also test for the impact of host behavior on the probability of short-term preference.  We find that more flexibility around the number of guests and the minimum stay increase the probability of short-term preference, while the reverse is true of cancellation policies (more strict equals higher short-term preference probability).  As a host, it appears to be best to do what you can to attract bookings and then make it difficult for those bookings to be cancelled.

@@

There are a number of limitations to this research.  There are few shared property characterstics between the two datasets, which make cross-estimating rents and rates difficult, especially from the short-term market.  Part of the difficulty in estimating potential short-term rates for long-term properties is that the short-term market appears to be heavily influenced by non-property specific factors such as the leasing policies (minimum stay, number of guests, etc.) themselves.  Future research should focus both on building better models to explain (and ultimately predict) short-term rates and occupancies as well as to understand the impact of policies such as strict cancellation or no pets on the short-term market.  It is likely that this work is being done proprietarily by Airbnb.com itself and other complementary firms such as AirDNA.com, our data provider.

@@

<!-- ```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='Logistic Model Results of short-term Preference', fig.lp='xxx'} -->
